A practical guide to Digital Micro-mirror Devices (DMDs) for wavefront shaping

### 2025-12-05 paper/main.tex

\definecolor{bg}{RGB}{255,249,227}
\definecolor{light_orange}{RGB}{255,193,134}
% \newcommand{\red}[1]{\textcolor{red}{#1}}
\newcommand{\red}[1]{#1}

@@

@@

Since the advent of adaptive optics, various technologies have been employed
to modulate the amplitude and/or phase of light.
Early adaptive optics devices, utilized in fields like microscopy and astronomy,
offer rapid modulation capable of compensating for the aberrations of optical systems
in real-time.
However, these devices are constrained by a limited number of actuators,
restricting their utility in complex media where a large number of degrees of freedom is essential.
Liquid Crystal Spatial Light Modulators (LC-SLMs),
which allow for the control of light phase across typically more than a million pixels,
have emerged as powerful tools for wavefront shaping in complex media
since the seminal work of A. Most and I. Vellekoop in the mid-2000s~\cite{Vellekoop2007focusing}.
Nonetheless, LC-SLMs are hampered by their slow response time,
permitting only a modulation speed ranging from a few Hz to \red{a few hundred} Hz.\\

@@

While other articles exist describing the various aspects of DMDs~\cite{Park2015properties, Scholes2019structured, Cox2021converting, Wang2023diffraction},
this tutorial aims to provide a guide for easily setting up a DMD for wavefront shaping applications in complex media.
In particular, we provide characterization and validation procedures that require minimal changes compared to typical wavefront shaping setups. 
\red{
More specifically, we place ourselves in a standard experiment where the goal is to shape the complex wavefront impinging on a complex medium and control or measure its output response. 
This is typically the case for a focusing experiment~\cite{Vellekoop2007focusing} or for measuring the transmission matrix of a complex medium~\cite{Popoff2010Measuring}.
}\\
We first introduce the diffraction properties of a DMD
and elaborate on how these could impact the system's efficiency.
We also furnish a straightforward criterion for selecting the appropriate DMD parameters for a specified excitation wavelength.
In the next section,
we delve into the aberration impacts brought about by the non-flatness of the DMD surface.
We demonstrate a simple process to characterize this effect and provide compensation solutions.
In the third segment,
we detail the influence of mechanical vibrations that are induced by the DMD's cooling system.
Lastly, we discuss how the thermalization of the DMD chip can potentially result in variations to the DMD response over time.

@@

@@

A significant distinction between liquid crystal modulators and DMDs
lies in the geometry of the pixel surface: 
The pixels of the DMD are not flat, but are mirrors that can be tilted 
in two different positions around a rotation axis
along the diagonal of the square shape of the pixels.
This difference gives rise to diffraction effects that can adversely affect
both modulation quality and system efficiency.
The impact of these diffraction effects is highly dependent on several factors:
the wavelength of illumination, the pixel pitch, and both the incident and outgoing angles.
Therefore, in conjunction with selecting an appropriate anti-reflection coating,
it is crucial to ensure that the pixel pitch is compatible with the specific configuration being used.
Texas Instruments offers chips with a variety of pixel pitches $d$,
ranging approximately from $5$ to $\sim25$ \textmu m~\cite{TI}.\\

To achieve a qualitative understanding of this issue,
we consider a 1D array of pixels as illustrated in Fig.~\ref{fig:grating_geom}.
Initially, let's assume that all pixels are in the same state and are illuminated by a plane wave originating from the far field.
Under these conditions, the pixelated modulator essentially functions as a grating,
with a period $d$ that is equivalent to the pixel pitch.
It is important to underscore that these modulators possess a hardware-limited fill factor,
typically around 90\%.
This translates to an effective active pixel size of $d' < d$.\\

In general, a grating gives rise to various diffraction orders
with different intensities
and angles $\theta_p$, as dictated by the grating equation:
$
 \text{sin}(\theta_p)+ \text{sin}(\alpha) = p\lambda/d = p \, \text{sin}(\theta_D),
$
where $\lambda$ is the wavelength of the light, $\alpha$ is the incident angle, 
$\theta_D = \arcsin \left(\lambda/S\right)$ is the angle of the first diffraction order,
and $p$ is an integer value denoting the orders of diffraction.
The intensity of the individual diffraction orders is influenced by
the response of a single pixel,
constituting the unit cell of the grating,
and that is governed by $d$, $d'$, and its rotation angle in the case of a DMD.
Importantly, we can decouple the effects of the grating's periodicity,
which influences the angles of the diffraction orders,
from the effects of the response of a single pixel,
which shapes the envelope of the angular response.\\

@@

For a case of normal incidence with an LC-SLM (Liquid Crystal Spatial Light Modulator),
we can assume that the response of a single pixel is uniform over its surface.
Similarly, in the scenario of a blazed grating,
such as a DMD, a linear phase slope is present on each pixel.
This is due to the tilt angle $\theta_B$ of the mirrors.
For an arbitrarily selected incidence angle $\alpha$,
a global phase slope is introduced.
This results in a trivial shift of the angular diffraction pattern by an angle $\alpha$.
The incident angle $\alpha$ has the effect of shifting the entire angular diffraction pattern
relative to the case of normal incidence,
while the blazed angle $\theta_B$
—the tilt angle of the micro-mirrors of the DMD—
only shifts the envelope by an angle of $2\theta_B$.
Whenever the fill factor approaches 100\%,
i.e. when $d \approx d'$,
the envelope for a flat grating achieves its maximum at $\theta_0 = -\alpha$;
this corresponds to the angle of the zeroth-order diffraction,
and the intensity is comparatively small for all other orders.
In these specific conditions, a singular diffraction order is visually perceived, corresponding to the optimum scenario.
The addition of a blazed angle $\theta_B$ results in both
a shift in the envelope and in the position of its maximum 
$\theta_0 = 2\theta_B - \alpha$.
In the general case,
this position may not align with a diffraction order anymore~\cite{Park2015properties}.
We provide in Appendix A.  a 1D computation of this effect.
A more accurate computation of the far field can be found  in ~\cite{Wang2023diffraction}.\\

We represent in Fig.~\ref{fig:gratings} the angular response of a flat grating and a blazed grating
for a 1D filling fraction of 95\% (correponding to a 2D filling fraction of $\approx 90$\%).
For a flat grating, the zero-th order contains most of the intensity,
the other orders being negligible in comparison.
For the blazed grating example shown,
we are in a situation close to the worst case scenario:
Two diffraction orders have a significant and comparable intensity,
and other orders also have non-negligible contributions.
In the optimal scenario, where the peak of the envelope corresponds to a diffraction order,
it results in a single diffraction order carrying the majority of the energy.
This state is achieved when the conditions of the blazed grating equation are fulfilled~\cite{Casini2014on}:

@@

\begin{figure}
  \centering
  \begin{subfigure}{0.49\textwidth}
    \centering
    \includegraphics[width = \textwidth]{images/gratings_flat.pdf}
    \label{fig:flat_grating}
  \end{subfigure}
  \begin{subfigure}{0.49\textwidth}
    \centering
    \includegraphics[width = \textwidth]{images/gratings_blazed.pdf}
    \label{fig:blazed_right}
  \end{subfigure}
  \caption{
    \textbf{Flat grating vs blazed grating.}
    Far field diffraction patterns for a 1D flat grating (left) and a 1D blazed grating (right)
    for an input angle of $\alpha = -20^\circ$,
    a filling fraction of 95\% (corresponding to a 2D filling fraction of $\approx 90$\%),
    a pixel tilt angle of $\theta_B = 5^\circ$,
    and a wavelength to pixel pitch ratio $\lambda/d=0.05$.
    Vertical lines represent the angles of the diffraction orders, peaked at $\theta_p$ which satisfy $\sin(\theta_p)+\sin(\alpha)=p\lambda/d$, 
    and the black dashed curve represents the amplitude of the field which peaks at $\theta_0=2\theta_B-\alpha$.
  }
  \label{fig:gratings}
\end{figure}

@@

To analyze more precisely the effect of diffraction in a DMD,
one needs to consider the 2D surface of the modulator.
We can establish a Cartesian coordinate system on the plane of the DMD,
with axes $x$ and $y$ aligned with the pixel sides
(refer to Fig.~\ref{fig:2d_geom}a).
The pixels are uniformly repeated along these axes.
However, a technical challenge arises
in that the axis of rotation of the pixels
aligns with the pixel diagonals,
resulting in a rotation by 45 degrees with respect to the $x$ and $y$ axes.
For the convenience of alignment and manipulation of the optical setup,
it is preferable to work with the incident and outgoing beams
which have the optical axis contained in the horizontal plane,
i.e. a plane parallel to the table surface.
A straightforward and common solution is to rotate the chip by 45 degrees
relative to the horizontal plane,
which aligns the pixel axis of rotation to be vertical.
This configuration is depicted in Fig.~\ref{fig:2d_geom}b.\\

@@

@@

We can quantify how close we are to the ideal case,
i.e., when satisfying the blazed this equation,
by defining a {\em blazed number} $\mu$ as introduced in~\cite{WFSnet_diffraction}:

@@

where $\mod{2}$ represents the modulo 2 operation.
$\mu$ is maximal and equals $1$ when the blazing equation is satisfied,
i.e. when one order of diffraction contains most of the energy,
and minimal when we are in the worst-case scenario,
i.e. when four orders of diffraction have a significant and equal intensity.\\

To demonstrate the effect, we first conduct a simulation of a DMD using Python
(refer to tutorial and code in~\cite{WFSnet_diffraction})
with two pixel pitches of
$d=7.6$\textmu m and $d=10.8$\textmu m,
under a coherent excitation at $\lambda=633$nm.
Fig.~\ref{fig:mu76} shows the estimated blazed number $\mu$
as a function of the angle of incidence $\alpha$ in the horizontal plane,
along with the far field diffraction pattern for two distinct incident angles. 
Fig.~\ref{fig:mu_2d} shows the blazed number as a function of both the incident angle 
and the wavelength for pixel pitches of $d=7.6$\textmu m and $d=10.8$\textmu m.
It should be noted that the efficiency of diffraction can be altered by adjusting
the angle of incidence $\alpha$.
However, its impact is relatively confined
within an acceptable angular range that aligns with experimental limitations
(i.e. for angles far from $\pm 90^\circ$).
Far-field patterns are centered around the maximum of the envelope
for an output angle $2\theta_B - \alpha$
(marked by a yellow cross).
We see that for small positive values of $\alpha$,
the pixel pitch of $d=10.8$\textmu m leads to a blazed number $\mu$
close to $1$.
It corresponds in the far field to having one bright order of diffraction
close to the maximum of the envelope.

\begin{figure}
  \includegraphics[width = \textwidth]{images/mu_1d.pdf}
  \caption{
    \textbf{Blazed number and far-field diffraction patterns.}
    Blazing number $\mu$ (Eq.(\ref{eq:blazed_number})) as a function of the incident angle $\alpha$  (top)
    for a pixel pitch of $d=7.6$\textmu m (a.)
    and $d=10.8$\textmu m (b.).
    Corresponding far-field diffraction patterns (bottom)
    for two incident angles $\alpha = -15^\circ$ and $\alpha = 0^\circ$.
    The red cross indicates the maximum of the envelope $\theta_\text{max} = 2\theta_B - \alpha$.
  }
  \label{fig:mu76}
\end{figure}

\begin{figure}
  \centering
  \includegraphics[width = \textwidth]{images/mu_2d.pdf}
  \caption{
    \textbf{Blazed number as a function of the wavelength and the incident angle.
    }
    Color map representation of the blazed number for a pixel pitch of $d=7.6$\textmu m (a.)
    and $d=10.8$\textmu m (b.).
  }
  \label{fig:mu_2d}
\end{figure}

To illustrate the link between the blaze number and the efficiency of the diffraction setup, 
we compute the ratio of the energy in the brightest diffraction spot to the total diffracted energy 
for different values of the wavelength, 
with fixed values for the pixel pitch (7.6 \textmu m) and incident angle (normal incidence).
We simulate the DMD with no gap between the pixels (i.e., a filling fraction of 100\%) 
and with 10 pixels in each direction, 
where all the mirrors of the pixels are in the same state. 
The results are shown in Fig.~\ref{fig:mu_vs_diff}.
We observe that the maxima (resp. minima) of the diffraction efficiency 
correspond to the maxima (resp. minima) of the blaze number. 
Note that the actual values of the minimum and maximum diffraction efficiency depend on the parameters of the DMD, such as pixel pitch, filling fraction, resolution, etc.\\

\begin{figure}
  \centering
  \includegraphics[width = \textwidth]{images/mu_vs_diffeff.pdf}
  \caption{
    \textbf{Blazed number vs diffraction efficiency.
    }
    Plot of the blazed number and the diffraction efficiency for a normal incidence and a pixel pitch of  7.6 \textmu m 
    as a function of the wavelength. 
    The diffraction efficiency is estimated by comparing the energy in a diffracted limited spot at the 
    location of the brightest diffraction order to the total diffracted energy.
  }
  \label{fig:mu_vs_diff}
\end{figure}

@@

\subsection{Modulation cross-talk}

@@

% \section*{Appendix A: 1D calculation of the DMD diffraction effect}
% \label{app:1D}

>>>>

\definecolor{bg}{RGB}{255,249,227}
\definecolor{light_orange}{RGB}{255,193,134}
\newcommand{\red}[1]{\textcolor{red}{#1}}
% \newcommand{\red}[1]{#1}

@@

% macro for red color

@@

Since the advent of adaptive optics, various technologies have been employed
to modulate the amplitude and/or phase of light.
Early adaptive optics devices, utilized in fields like microscopy and astronomy,
offer rapid modulation capable of compensating for the aberrations of optical systems
in real-time.
However, these devices are constrained by a limited number of actuators,
restricting their utility in complex media where a large number of degrees of freedom is essential.
Liquid Crystal Spatial Light Modulators (LC-SLMs),
which allow for the control of light phase across typically more than a million pixels,
have emerged as powerful tools for wavefront shaping in complex media
since the seminal work of A. \red{Mosk} and I. Vellekoop in the mid-2000s~\cite{Vellekoop2007focusing}.
Nonetheless, LC-SLMs are hampered by their slow response time,
permitting only a modulation speed ranging from a few Hz to \red{a few hundred} Hz.\\

@@

While other articles exist describing the various aspects of DMDs~\cite{Park2015properties, Scholes2019structured, Cox2021converting, Wang2023diffraction},
this tutorial aims to provide a guide for easily setting up a DMD for wavefront shaping applications in complex media.
In particular, we provide characterization and validation procedures that require minimal changes compared to typical wavefront shaping setups. 
More specifically, we place ourselves in a standard experiment where the goal is to shape the complex wavefront impinging on a complex medium and control or measure its output response. 
This is typically the case for a focusing experiment~\cite{Vellekoop2007focusing} or for measuring the transmission matrix of a complex medium~\cite{Popoff2010Measuring}. 
\red{
In such works,
complex modulation is usually achieved by encoding the optical phase
into the spatial displacement of binary fringes
displayed on the DMD,
followed by filtering high spatial frequencies in the Fourier plane~\cite{lee1979binary}.
Such a configuration permits multi-level complex modulation,
at the cost of a reduced spatial resolution.
The implementation and performance of such systems are discussed in detail in a separate tutorial~\cite{Gutierrez2024DMD}
and are not elaborated further here.
For the remainder of this paper, it will be assumed that the DMD is used
for complex modulation via such a method.
}
\\

We first introduce the diffraction properties of a DMD
and elaborate on how these could impact the system's efficiency.
We also furnish a straightforward criterion for selecting the appropriate DMD parameters for a specified excitation wavelength.
In the next section,
we delve into the aberration impacts brought about by the non-flatness of the DMD surface.
We demonstrate a simple process to characterize this effect and provide compensation solutions.
In the third segment,
we detail the influence of mechanical vibrations that are induced by the DMD's cooling system.
Lastly, we discuss how the thermalization of the DMD chip can potentially result in variations to the DMD response over time.

@@

A significant distinction between liquid crystal modulators and DMDs
lies in the geometry of the pixel surface: 
The pixels of the DMD are not flat, but are mirrors that can be tilted 
in two different positions around a rotation axis
along the diagonal of the square shape of the pixels.
This difference gives rise to diffraction effects that can adversely affect
both modulation quality and diffraction efficiency, 
\red{defined the fraction of the incident optical power that 
is redistributed into a single diffraction order.}.
The impact of these diffraction effects is highly dependent on several factors:
the wavelength of illumination, the pixel pitch, and both the incident and outgoing angles.
Therefore, in conjunction with selecting an appropriate anti-reflection coating,
it is crucial to ensure that the pixel pitch is compatible with the specific configuration being used.
Texas Instruments offers chips with a variety of pixel pitches $d$,
ranging approximately from $5$ to $\sim25$ \textmu m~\cite{TI}.\\

@@

To achieve a qualitative understanding of this issue,
we consider a 1D array of pixels as illustrated in Fig.~\ref{fig:grating_geom}.
Initially, let's assume that all pixels are in the same state and are illuminated by a plane wave originating from the far field.
Under these conditions, the pixelated modulator essentially functions as a grating,
with a period $d$ that is equivalent to the pixel pitch.
It is important to underscore that these modulators possess a hardware-limited fill factor,
typically around 90\%.
This translates to an effective active pixel size of $d' < d$.\\
\red{
We compare here two types of gratings:
Flat gratings, corresponding to LC-SLMs (Liquid Crystal Spatial Light Modulators),
in which all pixels are in the same state;
and blazed gratings, corresponding to DMDs,
in which all pixels are tilted by the same angle.
A grating gives rise to multiple diffraction orders,
whose amplitudes are modulated by a global envelope.
Importantly, these two effects can be decoupled.
The diffraction order angles $\theta_p$
are determined by the grating periodicity,
while the envelope, and in particular the position of its maximum $\theta_{\mathrm{envelope}}^{\max}$,
is governed by the structure of a single pixel,
which constitutes the unit cell of the grating.
}

\subsubsection{Diffraction orders\\}

In general, a grating gives rise to various diffraction orders
with different intensities
and angles $\theta_p$, as dictated by the grating equation:
\begin{equation}
    \text{sin}(\theta_p)+ \text{sin}(\alpha) = p\lambda/d = p \, \text{sin}(\theta_D),    
    \label{eq:1D_orders}
\end{equation}

where $\lambda$ is the wavelength of the light, $\alpha$ is the incident angle, 
$\theta_D = \arcsin \left(\lambda/d\right)$ is the angle of the first diffraction order,
and $p$ is an integer value denoting the orders of diffraction.
The intensity of the individual diffraction orders is influenced by
the response of a single pixel,
constituting the unit cell of the grating,
and that is governed by $d$, $d'$, and its rotation angle in the case of a DMD.
\red{
While this regime is not always relevant,
we can adopt the small-angle approximation to obtain
a qualitative picture of the effect of the incident angle $\alpha$
on the diffraction orders.
In this case, we have $\theta_p \approx p\lambda/d - \alpha$.
Thus, changing the incident angle effectively
rotates all diffraction orders by an angle $-\alpha$.
}\\
% Importantly, we can decouple the effects of the grating's periodicity,
% which influences the angles of the diffraction  \red{$\theta_p$},
% from the effects of the response of a single pixel,
% which shapes the envelope of the angular response.\\

@@

\subsubsection{Position of the maximum of the envelope\\}
% For a case of normal incidence with an LC-SLM (Liquid Crystal Spatial Light Modulator),
% we can assume that the response of a single pixel is uniform over its surface.
% Similarly, in the scenario of a blazed grating,
% such as a DMD, a linear phase slope is present on each pixel.
% This is due to the tilt angle $\theta_B$ of the mirrors.
\red{
The envelope of the diffracted field,
which sets the amplitude of each diffraction order,
is governed by the response of each unit cell, i.e. each pixel.
In a flat grating excited at normal incidence,
the envelope is peaked around the normal to the surface, i.e. $\theta_{envelope}^{max} = 0$.
In a blazed grating,
the tilt angle $\theta_B$ introduces a linear phase ramp
in the response of each individual pixel.
A flat grating can be seen as a particular case with $\theta_B = 0$.
The incidence angle $\alpha$ also adds
a global phase slope.
When the fill factor approaches 100\%,
i.e. when $d \approx d'$,
the cumulative effect leads to a shift of the envelope
by an angle $2\theta_B - \alpha$.
This results in a maximum of the envelope at
}

\begin{equation}
    \theta_{envelope}^{max}= 2\theta_B - \alpha
    \label{eq:1D_envelope}
\end{equation}

\red{
which reduces to $\theta_{envelope}^{max} = -\alpha$ in the case of a flat grating.
We detail the calculation of this effect in Appendix~A.
}

\subsubsection{Blazed grating condition\\}

\red{
In the case of a flat grating ($\theta_B = 0$), we see from Eqs.~\ref{eq:1D_orders} and~\ref{eq:1D_envelope}
that the position of the first order is always aligned with the maximum of the envelope
at the angle $-\alpha$,
when the filling fraction is close to 1.
In the general case $\theta_B \neq 0$,
$\theta_{envelope}^{max}$ does not, in general, coincide
with any diffraction order~\cite{Park2015properties, Wang2023diffraction}.
This leads to a redistribution of the energy of the incoming beam
into multiple diffraction orders,
resulting in a reduced diffraction efficiency
as well as cross-talk between different pixel states,
as studied in section~\ref{sub:crosstalk}.\\
}

\red{
The maximal diffraction efficiency is obtained 
when the maximum of the envelope matches one order of diffraction. 
This state is achieved when the conditions of the blazed grating equation are fulfilled~\cite{Casini2014on}, 
i.e. where there exist an integer $p$ so that we satisfy:
}

@@

\red{
In particular, in the Littrow configuration, 
i.e. for $\alpha = \theta_B$, we have $\theta_{max}^{envelope} = \alpha$; 
the diffraction and incidence angles are identical. 
The optimal condition is then satisfied for $\sin(\theta_B)={p\lambda}/{(2d)} = \sin(\theta_D)/2$.
}

% The later results in a trivial shift of the angular diffraction pattern by an angle $-\alpha$.
% The incident angle $\alpha$ has the effect of shifting the entire angular diffraction pattern
% relative to the case of normal incidence,
% while the blazed angle $\theta_B$
% —the tilt angle of the micro-mirrors of the DMD—
% only shifts the envelope by an angle of $2\theta_B$.
% Whenever the fill factor approaches 100\%,
% i.e. when $d \approx d'$,
% the envelope for a flat grating achieves its maximum at \red{$\theta_{envelope}^{max} = -\alpha$};
% this corresponds to the angle of the zeroth-order  \red{$\theta_{p=0}$},
% and the intensity is comparatively small for all other orders.
% In these specific conditions, a singular diffraction order is visually perceived, corresponding to the optimum scenario.
% The addition of a blazed angle $\theta_B$ results in both
% a shift in the envelope and in the position of its maximum 
% \red{$\theta_{envelope}^{max}= 2\theta_B - \alpha$. 
% The envelope is peaked at the angle corresponding to the reflection on a 
% unit cell, which consists in a mirror tilted with an angle $\theta_B$ 
% with respect to the surface of the grating.}
% \red{
% We show in Appendix A. that the angles of the diffraction orders
% are given by
% $
% \theta_p = \arcsin \left( 
%     p\,\lambda/d -sin(\alpha) \right)
% $,

% }

\subsubsection{Example\\}

\begin{figure}[ht]
  \centering
  \includegraphics[width = 0.75 \textwidth]{paper/images/Fig3bis.pdf}
  \caption{
    \textbf{1D grating example.}
    All angles are defined with respect to the normal of the surface of the the grating, 
    denoted by $\vec{n}_G$, and are taken as positive for clockwise rotations and negative for counter-clockwise rotations.
    The blazing angle $\theta_B$ is the angle between the normal to the micro-mirror $\vec{n}_M$ and $\vec{n}_G$.
    The angles of the diffraction orders $\theta_p$ are independent of $\theta_B$, while the maximum of the envelope is at $\theta_{max}^{envelope} = 2\theta_B -\alpha$.
  }
  \label{fig:grating_geom_example}
\end{figure}

\red{
We represent in Fig.~\ref{fig:grating_geom_example} an example of a geometry 
for a 1D blazed grating and 
in Fig.~\ref{fig:gratings} its angular response 
as well as the one corresponding to a flat grating with the same parameters.}
We take a 1D filling fraction of 95\% (correponding to a 2D filling fraction of $\approx 90$\%).
For a flat grating, the zero-th order contains most of the intensity,
the other orders being negligible in comparison.
For the blazed grating example shown,
we are in a situation close to the worst case scenario:
Two diffraction orders have a significant and comparable intensity,
and other orders also have non-negligible contributions.
In the optimal scenario, where the peak of the envelope corresponds to a diffraction order,
it results in a single diffraction order carrying the majority of the energy.
This state is achieved when the conditions of the blazed grating equation are fulfilled~\cite{Casini2014on}:
}

% \begin{equation}
%   \text{sin}(2\theta_B-\alpha) + \text{sin}(\alpha)
%   = 2 \,\text{sin}(\theta_B)  \,\text{cos}(\theta_B-\alpha)
%   = p\frac{\lambda}{d} \, .
%   \label{eq:blazed_eq}
% \end{equation}

% \red{
% In particular, in the Littrow configuration, i.e. for $\alpha = \theta_B$, 
% we have $\theta_{max}^{envelope} = \alpha$; 
% the diffraction and incidence angles are identical. 
% The optimal condition is then satisfied for 
% $\sin(\theta_B)={p\lambda}/{(2d)} = \sin(\theta_D)/2$.
%  }
% For the special case, where sin(theta_B)=p\lambda/d  that all the energy goes to the p order for \theta_b=\alpha. The intuition is that that for this wavelength a plane wave  accumulates phase of p*2*pi, so like a flat mirror.

\begin{figure}
  \centering
  \begin{subfigure}{0.49\textwidth}
    \centering
    \includegraphics[width = \textwidth]{images/flat_grating_lin_log.pdf}
    \label{fig:flat_grating}
  \end{subfigure}
  \begin{subfigure}{0.49\textwidth}
    \centering
    \includegraphics[width = \textwidth]{images/blazed_grating_lin_log.pdf}
    \label{fig:blazed_right}
  \end{subfigure}
  \caption{
    \textbf{Flat grating vs blazed grating.}
    Far field diffraction patterns for a 1D flat grating (left) and a 1D blazed grating (right)
    for an input angle of $\alpha = -20^\circ$,
    a filling fraction of 95\% (corresponding to a 2D filling fraction of $\approx 90$\%),
    a pixel tilt angle of $\theta_B = 5^\circ$,
    and a wavelength to pixel pitch ratio $\lambda/d=0.05$.
    Vertical lines represent the angles of the diffraction orders, peaked at $\theta_p$ which satisfy $\sin(\theta_p)+\sin(\alpha)=p\lambda/d$, 
    and the black dashed curve represents the amplitude of the field which peaks at $\theta_0=2\theta_B-\alpha$.
  }
  \label{fig:gratings}
\end{figure}

@@

% To analyze more precisely the effect of diffraction in a DMD,
% one needs to consider the 2D surface of the modulator.
% We can establish a Cartesian coordinate system on the plane of the DMD,
% with axes $x$ and $y$ aligned with the pixel sides
% (refer to Fig.~\ref{fig:2d_geom}a).
% The pixels are uniformly repeated along these axes.
% However, a technical challenge arises
% in that the axis of rotation of the pixels
% aligns with the pixel diagonals,
% resulting in a rotation by 45 degrees with respect to the $x$ and $y$ axes.
% For the convenience of alignment and manipulation of the optical setup,
% it is preferable to work with the incident and outgoing beams
% which have the optical axis contained in the horizontal plane,
% i.e. a plane parallel to the table surface.
% A straightforward and common solution is to rotate the chip by 45 degrees
% relative to the horizontal plane,
% which aligns the pixel axis of rotation to be vertical.
\subsubsection{Blazed grating condition\\}
\red{
A DMD consists of a 2D array of micro-mirrors.
In most DMDs, the rotation axis of the mirrors
is aligned with the pixel diagonals.
For convenience in the alignment and manipulation of the optical setup,
it is preferable to work with incident and outgoing beams
whose optical axes lie in the horizontal plane.
A straightforward and common solution is to rotate the chip by $45^\circ$
with respect to the horizontal plane,
which makes the pixels' rotation axis vertical.
This configuration is depicted in Fig.~\ref{fig:2d_geom}.
We show in Appendix B. that it leads to a new blazed grating equation that reads:
}\\

@@

% A more detailed description of the 2D system 
% and computation of its response 
% can be found in~\cite{Scholes2019structured}.
% However, we can find a simple condition for
% having most of the energy in one diffraction order 
% by looking at the effect of the tilts of the mirrors 
% when the axis is rotated by 45 degrees compared to 
% the axis of the pixel array (axes $x$ and $y$). 
% We consider a plane wave with an incoming angle $\alpha$ contained in the horizontal plane 
% (i.e. orthogonal to the rotation axis of the micro-mirrors).
% The effect of this angle is represented by a phase term of the form:

% \begin{equation}
% {\vec{k}_{in}\cdot\vec{r}} = {\frac{2\pi}{\lambda}\sin\alpha \frac{\left(x - y\right)}{\sqrt{2}}}
% \, ,
% \end{equation}

% with $\vec{k}_{in}$ the incoming wavevector and $\vec{r}$ the position in the modulator plane, 
% represented by the coordinate $(x,y)$.

% The grating condition is achieved when the output contributions are in phase 
% for the output angle corresponding to the specular reflection on the tilted mirrors. 
% It corresponds to the plot on the left in Fig.\ref{fig:gratings} for the 1D case.
% The phase contribution resulting from an output angle $2\theta_B-\alpha$, 
% corresponding to the specular reflection on each micro-mirror,
% reads:

% \begin{equation}
% \vec{k}_{out}\cdot\vec{r} 
% = {\frac{2\pi}{\lambda}\sin\left(2\theta_B - \alpha\right) \frac{\left(x - y\right)}{\sqrt{2}}}
% \, ,
% \end{equation}

% The total phase then reads:

% \begin{equation}
% \phi(x,y) = \vec{k}_{in}.\vec{r}  + \vec{k}_{out}.\vec{r} =  
% {\frac{2\pi}{\lambda}\sin\left(2\theta - \alpha\right) \frac{\left(x - y\right)}{\sqrt{2}}} +
% {\frac{2\pi}{\lambda}\sin\alpha \frac{\left(x - y\right)}{\sqrt{2}}} 
% \, .
% \end{equation}

% We see that $x$ and $y$ have the same effect.
% The grating condition is then achieved when two consecutive pixels, 
% separated by a distance $d$ in both the $x$ and $y$ directions, are in phase.
% I.e. for

% \begin{equation}
%   \phi(x,y) = \phi(x+d,y) = \phi(x,y+d) \, .
% \end{equation}

% The new blazed grating equation then reads~:

@@

\subsubsection{Blazed number\\}
We can quantify how close we are to the ideal case,
i.e., when satisfying the blazed equation,
by defining a {\em blazed number} $\mu$ as introduced in~\cite{WFSnet_diffraction}:

@@

where $\mod{2}$ represents the modulo 2 operation.
$\mu$ is maximal and equals $1$ when the blazing equation is satisfied,
i.e. when one order of diffraction contains most of the energy,
\red{ 
corresponding to one diffraction order $p$
being close to the maximum of the envelope
($\theta_p \approx \theta^{\mathrm{envelope}}_{\max}$).
$\mu$ is minimal in the worst-case scenario,
i.e. when four diffraction orders have a significant and equal intensity.
Note that, since the maximum of the envelope is no longer aligned
with the zeroth order, corresponding to the specular reflection on the micro-mirrors,
the value of the order $p$ that aligns with the envelope peak
when $\mu \approx 1$ can take any integer value,
and depends on the DMD parameters and on the angle of incidence $\alpha$.
}
\\

\subsubsection{Simulations\\}
To demonstrate the effect, we first conduct a simulation of a DMD using Python
(refer to tutorial and code in~\cite{WFSnet_diffraction})
with two pixel pitches of
$d=7.6$\textmu m and $d=10.8$\textmu m,
under a coherent excitation at $\lambda=633$nm.
Fig.~\ref{fig:mu76} shows the estimated blazed number $\mu$
as a function of the angle of incidence $\alpha$ in the horizontal plane,
along with the far field diffraction pattern for two distinct incident angles. 
Fig.~\ref{fig:mu_2d} shows the blazed number as a function of both the incident angle 
and the wavelength for pixel pitches of $d=7.6$\textmu m and $d=10.8$\textmu m.
It should be noted that the efficiency of diffraction,
\red{i.e. the fraction of the incident optical power that is diffracted into a given diffraction order 
relative to the incident power, }
can be altered by adjusting
the angle of incidence $\alpha$.
However, its impact is relatively confined
within an acceptable angular range that aligns with experimental limitations
(i.e. for angles far from $\pm 90^\circ$).
Far-field patterns are centered around the maximum of the envelope
for an output angle $2\theta_B - \alpha$
(marked by a \red{red} cross).
We see that for small positive values of $\alpha$,
the pixel pitch of $d=10.8$\textmu m leads to a blazed number $\mu$
close to $1$.
It corresponds in the far field to having one bright order of diffraction
close to the maximum of the envelope.

\begin{figure}
  \includegraphics[width = \textwidth]{images/mu_1d.pdf}
  \caption{
    \textbf{Blazed number and far-field diffraction patterns.}
    Blazing number $\mu$ (Eq.(\ref{eq:blazed_number})) as a function of the incident angle $\alpha$  (top)
    for a pixel pitch of $d=7.6$\textmu m (a.)
    and $d=10.8$\textmu m (b.).
    Corresponding far-field diffraction patterns (bottom)
    for two incident angles $\alpha = -15^\circ$ and $\alpha = 0^\circ$.
    The red cross indicates the maximum of the envelope $\theta^{envelope}_\text{max} = 2\theta_B - \alpha$
    \red{with $\theta_B = -12^\circ$}.
  }
  \label{fig:mu76}
\end{figure}

\begin{figure}
  \centering
  \includegraphics[width = \textwidth]{images/mu_2d.pdf}
  \caption{
    \textbf{Blazed number as a function of the wavelength and the incident angle.
    }
    Color map representation of the blazed number for a pixel pitch of $d=7.6$\textmu m (a.)
    and $d=10.8$\textmu m (b.) \red{for $\theta_B = -12^\circ$}.
  }
  \label{fig:mu_2d}
\end{figure}

To illustrate the link between the blaze number and the efficiency of the diffraction setup, 
we compute the ratio of the energy in the brightest diffraction spot to the total diffracted energy 
for different values of the wavelength, 
with fixed values for the pixel pitch (7.6 \textmu m) and incident angle (normal incidence).
We simulate the DMD with no gap between the pixels (i.e., a filling fraction of 100\%) 
and with 10 pixels in each direction, 
where all the mirrors of the pixels are in the same state. 
The results are shown in Fig.~\ref{fig:mu_vs_diff}.
We observe that the maxima (resp. minima) of the diffraction efficiency 
correspond to the maxima (resp. minima) of the blaze number. 
Note that the actual values of the minimum and maximum diffraction efficiency depend on the parameters of the DMD, such as pixel pitch, filling fraction, resolution, etc.
\red{
In a practical situation, pixels are not all in the same state:
the configuration is a mix of \textit{on} and \textit{off} states
corresponding to angles $\pm \theta_B$.
The diffraction efficiency then also depends
on the modulation scheme and on the specific parameters used for the modulation.
This topic is addressed in detail in a separate tutorial
focusing on complex modulation using binary amplitude modulators~\cite{Gutierrez2024DMD}.
}\\

\begin{figure}
  \centering
  \includegraphics[width = \textwidth]{images/mu_vs_diffeff.pdf}
  \caption{
    \textbf{Blazed number vs diffraction efficiency.
    }
    Plot of the blazed number and the diffraction efficiency for a normal incidence and a pixel pitch of  7.6 \textmu m 
    as a function of the wavelength. 
    The diffraction efficiency is estimated by comparing the energy in a \red{diffraction} limited spot at the 
    location of the brightest diffraction order to the total diffracted energy.
  }
  \label{fig:mu_vs_diff}
\end{figure}

@@

\subsection{Modulation cross-talk}
\label{sub:crosstalk}

@@

\section*{Appendix A: 1D calculation of the DMD diffraction effect}
\label{app:1D}
